Layered Formal Verification of an Industrial TCP/IP Stack

### 2021-06-02 SecDev21/paper.tex
moving the info part from results to intro

@@

Simplified contracts extracted from our library are reproduced in
\Cref{fig:functionorder}. If no error occurs, the function
\spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to
\spark{Remote_Ip_Addr} which is supposed to be
initialized when the function is called. \spark{Socket_Send} requires to be
called with a socket not null such that its field \spark{Remote_Ip_Addr} is
initialized. The only way to ensure that the precondition hold is to
call \spark{Socket_Connect} before \spark{Socket_Send}.
The tool GNATprove can statically find if a call to \spark{Socket_Send}
is not preceded by a call to \spark{Socket_Connect}.

@@

A common mistake is to forget to check the return codes after a call to a socket
API's function. For instance, the following code would be reject by SPARK:
\begin{lstlisting}[language=Ada,basicstyle=\small\ttfamily]
Socket_Connect(Sock, Ip_Addr, Port, Error);
Socket_Send(Sock, Data, Written, Flags, Error);
\end{lstlisting}
Indeed, if \spark{Socket_Connect} fails and \spark{Error} is not equal to
\spark{NO_ERROR}, \spark{Sock.S_Remote_Ip_Addr} might be non initialized, and
then the function \spark{Socket_Send} cannot be called because of its
precondition.

@@

\subsection{Verified parts}

For achieving this verification, 50 subprograms have been translated from C to
SPARK. This accounts for \np{2266} lines of code of which \np{1165} are logical
code to write and prove the specifications.

The specifications are fully proved with GNATprove in 9m54s on a 12 processors with 16GB RAM laptop.

% above: need time to prove automatically with GNATprove with the spec of your laptop (num of cores, amount of memory)

>>>>

In total, for this work, 50 subprograms have been translated from C to SPARK. This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g\.~ contracts) used to represent and prove the protocol's specification. The overall proving time with GNATprove is 9m and 54s on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.

@@

Simplified contracts extracted from our SPARK implementation are reproduced in \Cref{fig:functionorder}, to demonstrate how SPARK can be used to enforce the correct function calls' ordering. If no error occurs, the function \spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to \spark{Remote_Ip_Addr} which is supposed to be initialized when the function is called. \spark{Socket_Send} requires to be called with a socket not null such that its field \spark{Remote_Ip_Addr} is initialized. The only way to ensure that the precondition hold is to call \spark{Socket_Connect} before \spark{Socket_Send}. The tool GNATprove can statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.

@@

The following code demonstrates how SPARK will enforce to the check of the return codes of a function and allow execution to proceed only if no error code is returned:

\begin{lstlisting}[language=Ada,basicstyle=\small\ttfamily]
Socket_Connect(Sock, Ip_Addr, Port, Error); 
Socket_Send(Sock, Data, Written, Flags, Error);
\end{lstlisting}

In the above scenario the post-condition of  \spark{Socket_Connect} will set the value of \spark{Sock.S_Remote_Ip_Addr} if the function succesfully executes, and thus, \spark{Error} is not equal to \spark{NO_ERROR}. In the case an error occurs, \spark{Sock.S_Remote_Ip_Addr} won't be initialized, and thus, the function \spark{Socket_Send} cannot be called because its precondition wont be satisfied.

@@
